Appendix

### 2025-04-08 manuscript_rmarkdown/manuscript.Rmd
Update to text and tables.

@@

Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
we used $\kappa = 0.80$. Lookup tables and curves for critical statistic values for central normality and empirical central
normality tests are shown in Appendix E.

```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha = 0.95, n, \\kappa}$ of the (empirical) ",
  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
  "The critical test statistics for central normality test are determined using fully normal data, ",
  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
  "emp: empirical"
)

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=100$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(2, 4\right)$, resulting in a set of sequences with between $100$ and $10000$ elements.

@@

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.80, i}$, i.e. the mean of residual 
errors of the central 80% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.

@@

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
The \textit{optimal} column shows the optimal parameter values.
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
Lower loss indicates better transformation to central normality in the presence of outliers.}
\label{tab:optimal-weighting-parameters-box-cox}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.06 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.16 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.86 & 2.93 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.90 & 2.96 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 2.33 & ---  & ---       & ---  & 6.21 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.09 & 1.96 & $(0, 10]$ & 4.92 & 6.15 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.88 & 1.96 & $(0, 10]$ & 3.43 & 6.68 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 0.75 & ---  & ---       & ---  & 6.25 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.83 & 1.00 & $(0, 10]$ & 0.84 & 6.47 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.75 & 1.00 & $(0, 10]$ & 0.76 & 6.28 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
The \textit{optimal} column shows the optimal parameter values. 
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.
Lower loss indicates better transformation to central normality in the presence of outliers.}
\label{tab:optimal-weighting-parameters-yeo-johnson}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 6.19 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.81 & ---  & ---       & ---  & 3.13 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.00 & 0.95 & $(0, 1]$  & 0.87 & 2.73 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.14 & 0.95 & $(0, 1]$  & 0.91 & 2.80 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 3.32 & ---  & ---       & ---  & 5.50 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 0.30 & 1.96 & $(0, 10]$ & 5.91 & 5.78 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.53 & 1.96 & $(0, 10]$ & 5.08 & 5.16 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 1.05 & ---  & ---       & ---  & 7.51 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 1.04 & 1.00 & $(0, 10]$ & 1.08 & 7.61 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.17 & 1.00 & $(0, 10]$ & 1.54 & 8.80 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

## Assessing transformation using simulated data

@@

- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
  transformed using an inverse power transformation with randomly drawn 
  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
  were shifted into the positive domain, with minimum value $1 \cdot 10^{-4}$.

@@

This results in nine types of power transformation. Transformation parameters
were optimised for each sequence. Subsequently the sum of residual errors 
and sum of residual errors of the central portion ($\kappa = 0.70$) were computed
for each sequence after transformation. Then, each method was ranked according 
to the sum of residual errors (data without outliers) or the sum of residual
errors of the central portion (data with outliers). The average rank of each
transformation method was computed over all 10000 sequences in each dataset.
For Yeo-Johnson transformations average ranks are shown in Table \ref{tab:comparison_methods_simulation_yeo_johnson}.

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_yeo_johnson}
\begin{tabular}{l | l r r r r r r}

@@

none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\

>>>>

```{r include = FALSE, eval = FALSE}
# This creates and sets the lookup table for critical test statistic.
load(file = "./R/sysdata.rda")

central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.80,
  for_manuscript = FALSE
)

empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.80,
  for_manuscript = FALSE,
  with_outliers = TRUE
)

@@

Since the (empirical) central normality tests assesses whether the central part of a sequence is normally distributed,
we used $\kappa = 0.80$. Critical statistic values for central normality and empirical central
normality tests are shown are shown in Table \ref{tab:central-normality-critical-statistic} and Table 
\ref{tab:empirical-central-normality-critical-statistic}, respectively.

```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$ of the (empirical) ",
  "central normality test as function of $n$ for several values of central portion $\\kappa$. ",
  "The critical test statistics for central normality test are determined using fully normal data, ",
  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
  "emp: empirical"
)

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.

@@

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised a composite loss $L = \sum_{i=1}^{m_d} L_{\text{cn},i} + 0.1 \sum_{i=1}^{m_d} L_{\lambda,i}$.
The composite loss consisted of two components: a loss term $L_{\text{cn},i} = \tau_{n_i, \kappa = 0.80,i}$, i.e. the mean of residual 
errors of the central 80% of elements of each sequence $\mathbf{X}_i$, which aimed at optimising central normality; 
and a loss term $L_{\lambda,i} = \max\left(0.0,|\lambda_{0,i} - \lambda_{i}| - \xi \right)$,
with $\lambda_{0,i}$ and $\lambda_{i}$ transformation parameters found
for sequence $\mathbf{X}_i$ prior to and after adding outliers, respectively, 
and tolerance parameter $\xi = 0.5$ for Box-Cox and $\xi = 0.3$ for Yeo-Johnson power transformations.
This second term aimed to prevent solutions that provide small improvements in central normality
at the cost of a poor fit of the tails of sequences.

@@

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Box-Cox power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column. 
The \textit{optimal} column shows the optimal parameter values.
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
\label{tab:optimal-weighting-parameters-box-cox}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 49.6 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.80 & ---  & ---       & ---  & 38.5 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.01 & 1.00 & $(0, 1]$  & 0.86 & 43.7 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.00 & 1.00 & $(0, 1]$  & 0.90 & 42.7 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 2.39 & ---  & ---       & ---  & 52.2 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 2.39 & 2.40 & $(0, 10]$ & 4.92 & 52.7 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 1.07 & 3.63 & $(0, 10]$ & 3.43 & 57.4 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 0.83 & ---  & ---       & ---  & 53.1 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.80 & 0.81 & $(0, 10]$ & 0.84 & 54.4 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 0.76 & 0.77 & $(0, 10]$ & 0.76 & 53.2 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Optimal weighting parameters and corresponding loss for location- and scale-invariant Yeo-Johnson power transformations.
$p^{*}$ indicates use of the empirical distribution of feature values, $z$ the z-score of the transformed feature values,
and $r$ the residual error between the z-score of transformed feature values and the expected z-score according to the normal distribution. 
The \textit{initial} column shows the starting parameter value for the optimisation process, with the corresponding boundary values in the \textit{limits} column.
The \textit{optimal} column shows the optimal parameter values. 
The \textit{loss} column shows the loss achieved by each method, under optimised parameters.}
\label{tab:optimal-weighting-parameters-yeo-johnson}
\begin{tabular}{l r r r r r r r}

@@

\midrule
non-robust               & ---  & ---       & ---  & ---  & ---       & ---  & 42.1 \\
$p^{*}$ (step)           & 0.80 & $(0, 1]$  & 0.78 & ---  & ---       & ---  & 35.1 \\
$p^{*}$ (triangle)       & 0.80 & $(0, 1]$  & 0.20 & 0.95 & $(0, 1]$  & 1.00 & 34.0 \\
$p^{*}$ (tapered cosine) & 0.80 & $(0, 1]$  & 0.54 & 0.95 & $(0, 1]$  & 1.00 & 32.6 \\
$z$ (step)               & 1.28 & $(0, 10]$ & 2.32 & ---  & ---       & ---  & 43.0 \\
$z$ (triangle)           & 1.28 & $(0, 10]$ & 1.28 & 1.96 & $(0, 10]$ & 3.43 & 49.7 \\
$z$ (tapered cosine)     & 1.28 & $(0, 10]$ & 0.41 & 1.96 & $(0, 10]$ & 3.75 & 51.3 \\
$r$ (step)               & 0.50 & $(0, 10]$ & 0.92 & ---  & ---       & ---  & 65.4 \\
$r$ (triangle)           & 0.50 & $(0, 10]$ & 0.87 & 1.00 & $(0, 10]$ & 0.87 & 66.0 \\
$r$ (tapered cosine)     & 0.50 & $(0, 10]$ & 1.06 & 1.00 & $(0, 10]$ & 1.07 & 67.1 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

## Assessing transformations using simulated data

@@

- A \textit{clean} dataset with each sequence drawn from $\mathcal{N}(0, 1)$ and
  transformed using an inverse power transformation with randomly drawn 
  transformation parameter $\lambda \sim U\left(0.00, 2.00 \right)$. 
  $\left(\phi_{\text{BC}}^\lambda\right)^{-1}$ and $\left(\phi_{\text{YJ}}^\lambda\right)^{-1}$
  were used as inverse transformations for assessing Box-Cox and Yeo-Johnson transformations,
  respectively. Prior to inverse transformation, sequences for Box-Cox transformations
  were shifted into the positive domain, with minimum value $1$.

@@

This results in nine types of power transformation. Transformation parameters
were optimised for each sequence. Subsequently the sum of residual errors 
and sum of residual errors of the central portion ($\kappa = 0.80$) were computed
for each sequence after transformation. Then, each method was ranked according 
to the sum of residual errors (data without outliers) or the sum of residual
errors of the central portion (data with outliers). The average rank of each
transformation method was computed over all 10000 sequences in each dataset.
Average ranks for Yeo-Johnson transformations are shown in Table
\ref{tab:comparison_methods_simulation_yeo_johnson}.
Location and shift-invariant Yeo-Johnson transformation ranked best for
all datasets without outliers. The robust variant ranked best in dirty and shifted
datasets with outliers, but not in the clean dataset.
Results for Box-Cox transformations are shown in Table \ref{tab:comparison_methods_simulations_box_cox}.

@@

\begin{table}
\begin{center}
\caption{
Comparison of average rank between Yeo-Johnson transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_yeo_johnson}
\begin{tabular}{l | l r r r r r r}

@@

none                                  & &         8.51  &         7.60  &         7.33  &         6.15  &         7.44  &         6.60 \\
conventional                          & &         3.70  &         5.70  &         3.87  &         7.06  &         6.49  &         5.93 \\
conventional (z-score norm.)          & &         4.72  &         5.68  &         6.38  &         5.01  &         5.43  &         4.87 \\
conventional (robust scaling)         & &         4.06  &         6.77  &         5.37  &         5.75  &         4.41  &         5.63 \\
Raymaekers-Rousseeuw                  & &         4.50  &         3.20  &         4.01  &         4.13  &         6.27  &         5.65 \\
Raymaekers-Rousseeuw (z-score norm.)  & &         5.27  & \textbf{2.66} &         6.23  &         3.59  &         5.28  &         3.48 \\
Raymaekers-Rousseeuw (robust scaling) & &         4.70  &         2.95  &         5.32  &         3.77  &         4.36  &         3.66 \\
invariant                             & & \textbf{3.13} &         6.77  & \textbf{3.22} &         6.12  & \textbf{2.64} &         6.03 \\
robust invariant                      & &         6.41  &         3.66  &         3.27  & \textbf{3.41} &         2.68  & \textbf{3.10} \\

### 2025-04-08 manuscript_rmarkdown/manuscript_appendix.Rmd
Update to text and tables.

@@

@@

# Appendix E: Empirical central normality test

```{r include = FALSE, eval = FALSE}
# This writes the lookup table for critical test statistic.
load(file = "./R/sysdata.rda")

central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.80,
  for_manuscript = FALSE
)

empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.80,
  for_manuscript = FALSE,
  with_outliers = TRUE
)

usethis::use_data(
  central_normality_lookup_table,
  empirical_central_normality_lookup_table,
  two_sided_function_parameters, 
  internal = TRUE, 
  overwrite = TRUE)
```

```{=latex}
\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.70,
  for_manuscript = TRUE,
  with_outliers = TRUE
)

```{=latex}
\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.70}$ of the (empirical) ",
  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
  "The critical test statistics for central normality test are determined using fully normal data, ",
  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
  "emp: empirical"
)

>>>>

# Appendix D: Central normality test and empirical central normality test critical values

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
\textbf{central normality} at $\kappa = 0.80$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\

\midrule
    5 & 210.79 & 95.75 & 66.27 & 49.69 & 35.93 & 27.60 & 19.29 & 13.41 & 11.00 \\
   10 &  48.04 & 31.44 & 26.96 & 23.91 & 21.11 & 18.16 & 13.66 & 10.14 &  8.67 \\
   20 &  29.13 & 22.13 & 19.75 & 17.94 & 16.00 & 13.81 & 10.54 &  8.18 &  7.18 \\
   50 &  17.31 & 14.01 & 12.63 & 11.48 & 10.28 &  9.00 &  7.00 &  5.54 &  4.94 \\
  100 &  12.19 & 10.04 &  9.02 &  8.21 &  7.38 &  6.47 &  5.04 &  3.99 &  3.56 \\
  200 &   8.68 &  7.11 &  6.40 &  5.83 &  5.22 &  4.57 &  3.59 &  2.86 &  2.56 \\ 
  500 &   5.61 &  4.54 &  4.08 &  3.70 &  3.33 &  2.91 &  2.28 &  1.82 &  1.63 \\
 1000 &   3.93 &  3.19 &  2.87 &  2.60 &  2.33 &  2.05 &  1.61 &  1.29 &  1.15 \\
 2000 &   2.71 &  2.24 &  2.02 &  1.84 &  1.66 &  1.46 &  1.14 &  0.91 &  0.82 \\
 5000 &   1.73 &  1.42 &  1.28 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
10000 &   1.22 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.37 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.80,
  for_manuscript = TRUE,
  with_outliers = TRUE
)

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.001 & 0.01 & 0.025 & 0.05 & 0.1 & 0.2 & 0.5 & 0.8 & 0.9 \\

\midrule
    5 & 159.27 & 66.78 & 57.06 & 50.55 & 43.42 & 34.89 & 22.28 & 14.92 & 12.10 \\
   10 &  41.82 & 30.82 & 27.01 & 24.31 & 21.58 & 18.53 & 13.82 & 10.23 &  8.75 \\
   20 &  29.44 & 23.65 & 21.23 & 19.16 & 17.01 & 14.67 & 11.10 &  8.42 &  7.35 \\
   50 &  18.95 & 15.56 & 13.96 & 12.74 & 11.35 &  9.82 &  7.51 &  5.83 &  5.16 \\ 
  100 &  13.49 & 11.00 &  9.91 &  8.98 &  8.03 &  7.04 &  5.43 &  4.25 &  3.76 \\
  200 &   9.65 &  7.99 &  7.22 &  6.64 &  5.96 &  5.21 &  4.05 &  3.17 &  2.81 \\
  500 &   6.37 &  5.32 &  4.82 &  4.44 &  4.03 &  3.57 &  2.79 &  2.18 &  1.92 \\
 1000 &   4.64 &  3.95 &  3.65 &  3.39 &  3.11 &  2.78 &  2.20 &  1.73 &  1.53 \\
 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.89 &  1.50 &  1.33 \\
 5000 &   2.78 &  2.48 &  2.36 &  2.24 &  2.11 &  1.95 &  1.66 &  1.39 &  1.25 \\
10000 &   2.39 &  2.18 &  2.08 &  2.00 &  1.91 &  1.81 &  1.59 &  1.39 &  1.28 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80}$ of the (empirical) ",
  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
  "The critical test statistics for central normality test are determined using fully normal data, ",
  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
  "emp: empirical"
)

@@

```
\begin{table}
\begin{center}
\caption{
Comparison of average rank between Box-Cox transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.80$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Box-Cox transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
If necessary, each sequence was shifted so that every element had a strictly positive value. 
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations_box_cox}
\begin{tabular}{l | l r r r r r r}

\toprule
& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
transformation & outliers: & no & yes & no & yes & no & yes \\

none                                  & &         7.65  &         6.96  &         7.28  &         6.58  &         7.45  &         6.78 \\
conventional                          & &         3.10  &         5.98  &         5.42  &         6.16  &         6.49  &         6.10 \\
conventional (z-score norm.)          & &         4.19  &         6.28  &         4.85  &         6.33  &         4.38  &         5.85 \\
conventional (robust scaling)         & &         4.20  &         6.28  &         4.78  &         6.33  &         4.30  &         5.86 \\
Raymaekers-Rousseeuw                  & &         4.46  &         3.55  &         5.39  &         3.69  &         6.29  &         5.84 \\ 
Raymaekers-Rousseeuw (z-score norm.)  & &         6.21  &         3.83  &         4.95  &         3.85  &         4.51  &         3.50 \\ 
Raymaekers-Rousseeuw (robust scaling) & &         6.12  &         3.82  &         4.97  &         3.85  &         4.52  &         3.50 \\
invariant                             & & \textbf{2.99} &         5.55  & \textbf{3.65} &         6.01  & \textbf{3.52} &         5.57 \\
robust invariant                      & &         6.09  & \textbf{2.75} &         3.71  & \textbf{2.20} &         3.55  & \textbf{2.00} \\

\bottomrule
\end{tabular}
\end{center}
\end{table}

@@
